### 2025-11-14 write-ups/manuscript/kid-tangrams-ms.Rmd
examples for expt 3, latex adventures

@@

In an iterated reference game using natural language to communicate, 4-8-year-old children succeeded when playing with a parent using a simple, child-friendly tablet-based task [@leung2024]. In this task, even 4-year-olds had an initial accuracy above 80% on two-way choice, which rose to above 90% in later repetitions [@leung2024]. Together, the recent work on both iterated and non-iterated reference games with young children provides evidence of young children’s ability to communicate with each other and with adults under the right conditions. The late preschool years are a time of rapid development for children as a number of communicative skills are improving between the ages of 3 and 6 [TODO maybe more citations, or re-citations].

Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, here we revisit the question of child-child referential communication. In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm. Across three studies including a total of ZZ pairs of 4-5-year-old preschool children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.

@@

4 and 5-year-old children were recruited from a university preschool during the school day. Children played with another child from the same class. Experiment 1 was conducted between June and August 2023. Pairs of children were included in analyses if they completed at least 8 of the 12 critical trials. We had 19 games that completed 12 critical trials, and 1 game that completed 11 critical trials. Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months. An additional XX games were excluded for being pilot runs or not completing enough trials.

@@

```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Length of description produced by the teller each trial. Grey dots are individual data points, colored dots are per trial means with bootstrapped 95% CIs. \\label{length}" }
joint_data |> 
  ggplot(aes(x = trial+1, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
    geom_vline(xintercept = 4.5, lty="dashed")+
    geom_vline(xintercept = 8.5, lty="dashed")+
    geom_vline(xintercept = 12.5, lty="dashed")+
  scale_color_manual(values = time_palette) +
  geom_smooth(method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  facet_grid(str_c("Expt ",expt) ~ .) +
  theme(
    legend.position = "right",
    panel.grid = element_blank(),
    strip.background = element_blank()
  )
```

@@

TODO this description metric is just the length of the description, taking out repetition/rephrasing from the child. We could also include the robustness check which is the length of all the (on-topic) teller-guesser interaction (including repeating/rephrasing, questions, I don't knows, etc), which is also very flat.

[TODO maybe move this paragraph] Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example} (these specific examples are from Experiment 2, but both experiments had similar distributions of descriptions).

@@

```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Semantic similarity between pairs of descriptions from different sources. Dots are means and lines are bootstrapped 95% CIs.  \\label{sim}" }

@@

TODO consider dropping this because it's not that strong an analysis?

@@

```{=latex}
\begin{table}
	\caption{Example descriptions children successfully used to identify different target images in Experiment 2. TODO UPDATE EXAMPLES TO HAVE SOME FROM EXPT 3
}
	\label{example}
	\begin{tabular}[t]{p{16em}p{4em}}
		\hline
		%hold
		$\bullet$ person & 		\multirow{6}{*}{\includegraphics[width=.5in]{I1.jpg}}\\

		$\bullet$ a person holding a sandwich &\\
		$\bullet$ a people carrying a box of dirt &\\
		$\bullet$ a monster &\\
		$\bullet$ someone holding a plate and giving it to a restaurant and has watermelon &\\
		\hline
		$\bullet$ vampire & \multirow{7}{*}{\includegraphics[width=.5in]{E1.jpg}}\\
		$\bullet$ hopping & \\
		$\bullet$ a person flying &\\
		$\bullet$ a person &\\
		$\bullet$ a kite &\\
		$\bullet$ a triangle with a head on it with feet &\\
				\multicolumn{2}{l}{$\bullet$  somebody skydiving, not in the airplane}\\
		\hline
		$\bullet$ racecar & \multirow{7}{*}{\includegraphics[width=.5in]{D1.jpg}}\\
		$\bullet$ airplane &\\
		$\bullet$ alligator &\\
		$\bullet$ a person fell down &\\
		$\bullet$ a boat &\\
		$\bullet$ a person that's in a race car that has one triangle and two triangles &\\
		\hline

		$\bullet$ person & \multirow{4}{*}{\includegraphics[width=.5in]{B1.jpg}}\\
		$\bullet$ a person walking &\\
		$\bullet$ a person looking down &\\
		$\bullet$ a people, but it doesn't have any arms& \\
		\hline
	\end{tabular}
\end{table}

@@

### Participants
Experiment 2 was run between March and August of 2024, at the same preschool as Experiment 1. No children participated in both experiments. 30 pairs of children completed all 16 critical trials, and 1 pair of children completed 10 critical trials. Our target age range was 4 and 5-year-olds, but one older 3-year-old was unintentionally included. Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months. An additional XX games were excluded for being pilot runs or not completing enough trials.

@@

As Experiment 2 had 4 blocks, we examined whether descriptions were converging semantically toward the final description. We compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity.^[`r sim_to_last_form |> form()`] Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`). Descriptions were marginally more similar if they came from the same child, which is expected as a sign of internal consistency TODO rephrase (`r sim_to_last_2_results |> stats_text(3,3)`). Although over time descriptions did get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant (`r sim_to_next_2_results |> stats_text(2,3)`).

@@

```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiments 2 and 3.  Heavy dots are means with bootstrapped 95% CIs; light dots are individual values. \\label{tolast}" }
joint_similarities |> 
  filter(game1==game2) |> 
  filter(later==3) |> 
  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>left_join(labels) |> 
  ggplot(aes(earlier+1, sim, color = img_source)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
    stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2), geom="line") +
    geom_smooth(color="black", method="lm")+
  facet_grid(str_c("Expt ", expt)~same_speaker) +
  scale_x_continuous(breaks=c(1,2,3), labels=c("1", "2", "3"))+
  labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
scale_color_brewer(palette="Dark2") +
  theme(
    legend.position = "right",
    legend.text = element_markdown(color = "black", size = 11),
    panel.grid = element_blank(),
    strip.background = element_blank()
  )

@@

## Methods
As Experiment 2 was very similar to Experiment 2, we focus on the changes made compared to Experiment 2. Experiment 3 was pre-registered at TODO [anonymized link].

@@

### Participants
Experiment 3 was run between XX and YY, at the same preschool as Experiments 1 and 2. Because the stimuli changed, we allowed children who had participated in a prior experiment to participate again, and ZZ participants in experiment 2 were also in experiment 3. TODO pairs of children completed all 16 critical trials, and TODO others completed ... . Our target age range was 4 and 5-year-olds. Of the TODO children, QQ were girls, and the children had a median age of ZZ months, and a range of XX-YY months. An additional XX games were excluded for being pilot runs, including repeat participants, not completing enough trials, or including a 3-year-old participant.

### Materials
A different set of critical images were used (see Figure TODO). These were chosen from a set of commonly used tangram images (CITATIONS), to be fairly distinct from each other. We also switched the stimuli for the 3rd and 4th practice trials to be more abstract, so children would not be surprised by the critical stimuli not having convention names (TODO figure).

### Procedure
The procedure was the same as for Experiment 2. The bubble popping practice was modified so that both rounds of bubbles were colored dots, because children in Experiment 2 sometimes did not identify the smiley faces as more bubbles to pop.

@@

description_mod_meta <- read_rds(here(mod_results, "description_meta.rds"))

@@

As the experiments were all similar to one another, we re-ran models pooling the data across the three experiments, with experiment number as a random effect. Pooling across the three experiments, children's accuracy was above chance (OR: `r stats_text(acc_mod_meta,1)`) and accuracy numerically increased over the course of the game, although the credible interval included 0 (OR of one trial later: `r stats_text(acc_mod_meta, 2)`).

TODO decide whether to include speed here again.

Descriptions produced by tellers averaged `r stats_text(description_mod_meta, 1)` words on the first trial, and description length was relatively stable over time (change of `r stats_text(description_mod_meta, 2)` words per trial).
Utterances were more similar if they came from the same partnership (increase in cosine similarity: `r sim_across_between_mod_meta |> stats_text(2,3)`) and were more similar still if they came from the same person within the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).

For the other semantic analyses, we pooled across experiments 2 and 3 which had 4 blocks. In the combined model, we found that the similarity the last block utterance increased across blocks (`r stats_text(sim_to_last_mod_meta, 2,3)`) and was higher when the same child produced both utterances (`r stats_text(sim_to_last_mod_meta, 3,3)`). In a combined model, similarity to the next block utterance also increased across blocks (`r stats_text(sim_to_next_mod_meta, 2,3)`) and was higher when the same child produced both utterances (`r stats_text(sim_to_next_mod_meta, 3,3)`). Across the two experiments, utterances from different pairs to the same stimulus tended to decrease in similarity across blocks (`r stats_text(sim_across_mod_meta, 2,3)`).

@@

We might expect that whether a description was successful influences whether the same description, or a variant of it, is employed in future rounds. Intuitively, successful descriptions can be copied and built upon, while unsuccessful descriptions should be replaced by a fresh attempt. To test whether accuracy is predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar to the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller. This pattern of results is consistent with the expectation that children are more likely to stick to their own description or repeat the other child's description if it was previously successful. <!--Numerically, the size of the boost for being correct was similar to the size of the boost from coming from the same person `r stats_text(sim_acc_to_next_mod_meta, 6,3)`. -->

@@

Children in the late preschooler years are undergoing a time of rapidly developing linguistic and cognitive skills, so children may show age-based differences in how well they complete tasks that use language or executive functioning [citation needed]. Working memory and other executive capacities are developing across childhood [@diamond2013]. Additionally, there may be gender differences at this age. Young girls have on average an advantage in early linguistic skills, showing faster earlier vocabulary growth that boys [@marchman; @rinaldi2023], and this linguistic advantage continues until .... TODO

@@

```{r age-words, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=2, fig.cap = "Number of words of description produced by the teller across all experiments, graphed by the age of teller in that game, and colored by teller gender. Each dot is a game, and linear fits are overlayed. \\label{agewords}" }
joint_data |>  ggplot(aes(x = speaker_age, y = words, color = speaker_gender)) +
  geom_smooth(method = "lm", show.legend=F)+
  geom_jitter(data=joint_data |> group_by(speaker_age, speaker, expt, game, speaker_gender) |> summarize(words=mean(words, na.rm=T)))+
  scale_color_manual(values=c("#268BD2FF", "#DC322FFF"))+
  labs(x="Teller Age (months)", y="Words", color="Teller gender")
```

@@

The effects of children's age and gender on their performance are shown in Figure \ref{ageacc}. We ran a mixed effects logistic regression, where we considered possible effects of the age of each child (centered at mean age of all the children who participated) and the gender of each child on the accuracy of their selections^[`r form(acc_demog_form)`]. Accuracy was higher for older children (OR for one month older speaker: `r stats_text(acc_demog_meta, 3,2)`, OR for one month older listener: `r stats_text(acc_demog_meta, 2,2)`). Accuracy was also higher when the listener was a girl (OR: `r stats_text(acc_demog_meta, 4,2)`), but was not credibly higher when the speaker was a girl (OR: `r stats_text(acc_demog_meta, 5,2)`).

TODO be consistent about teller/guesser versus speaker/listener

We also tested whether children's age or gender affected how many words they produced in their descriptions (Figure \ref{agewords}). We ran a mixed effects linear regression, where we considered the ages and genders of both children as possible predictors of description length^[`r form(description_demog_form)`]. Older children tended to produce longer descriptions (effect of one month older `r stats_text(description_demog_meta,3,2)`), but there was no effect of age of listener (`r stats_text(description_demog_meta, 2,2)`). For gender, there was no effect of teller gender on description length (girls produced `r stats_text(description_demog_meta, 5,2)` words compared to boys), but tellers produced longer descriptions for boy guessers (`r stats_text(description_demog_meta, 4,2)`). There were not substantial effects of trial order or interactions between demographics and trial order.

It is curious that listener gender but not speaker gender affected description length, but for this task, it is unclear how description length would be expected to map onto linguistic sophistication for this task and demographic. At one end, longer descriptions may provide more information than short non-specific labels such as "person" or "diamond". On the other, succinct descriptions that reflect distinguishing features ("mountain", "man with backpack") are better than long descriptions about the component shapes ("there's a triangle and then a square and another square...").

For accuracy, the age and gender effects could be reflective of any of a number of developing capacities, including linguistic sophistication, improving executive functioning and impulse control, and better perspective taking and theory of mind.

# Qualitative variation in interaction structure

@@

@@

* maybe cut this section
* do want to make the point that there's really wide variability (consistent with older kids and also adults for that matter) -- some pairs one kid does the describing including of options and the other chooses

* may want to be clear that structuring the interaction in an effective way is a skill that not all the kiddos have -- some relied more on the experimenter saying whose turn and needed more prompting with "do you want to give more hints" or "do you want to ask a question". whereas other groups rapidly produced descriptions and then either clicked or asked for (and got) clarification (either completely on their own, or after the option was questions offered as a possibility). 
that said, even when children were struggling, some of their attempts such as gesturing or turning ipads or asking (can you show me) are indicative of attempts at conversational monitoring and solving the communication problem (albeit not in ways that followed the "rules" of the game)
where we see what looks like failures it may be hard to tell what part isn't working (is it knowing what i know, or what other knows, or cognitive control, or producing questions, or sampling)

Prominent early studies claimed that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966]. However, more recent developmental work has found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low. Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents. Here, we revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.

Across 3 experiments and TODO pairs of 4 and 5-year-old children, we tested how well children could produce referential expressions that allowed their partner to find a matching abstract shape. Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (TODO%), indicating that children were generally able to produce adequate descriptions. Additionally, children's utterances showed signs of converging toward conceptual pacts. While this task was substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.

Unlike adults, children did not display an increase in accuracy or a shortening of referential expressions over the course of the game. Still, our findings show that descriptions became increasingly similar to descriptions in the last block and that successful utterances were more similar to future utterances, suggesting that children are adapting their descriptions as the game unfolds. These null findings are likely a result of initially high accuracy in the first block and initially short utterances that leave little room for reduction.

>>>>

```{r}
# for use in abstract 
pairs <- joint_data |> distinct(game, expt) |> nrow()
# eventually expect 81

pct <- joint_data |> summarize(m=round(mean(correct)*100)) |> pluck()
```

@@

In an iterated reference game using natural language to communicate, 4-8-year-old children succeeded when playing with a parent using a simple, child-friendly tablet-based task [@leung2024]. In this task, even 4-year-olds had an initial accuracy above 80% on two-way choice, which rose to above 90% in later repetitions [@leung2024]. Together, the recent work on both iterated and non-iterated reference games with young children provides evidence of young children’s ability to communicate with each other and with adults under the right conditions. The late preschool years are a time of rapid development for children as a number of communicative skills are improving between the ages of 3 and 6 years old.

Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, here we revisit the question of child-child referential communication. In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm. Across three studies including a total of `r pairs` pairs of 4-5-year-old preschool children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.

@@

4 and 5-year-old children were recruited from a university preschool during the school day. Children played with another child from the same class. Experiment 1 was conducted between June and August 2023. Pairs of children were included in analyses if they completed at least 8 of the 12 critical trials. We had 19 games that completed 12 critical trials, and 1 game that completed 11 critical trials. Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months. An additional 8 games were excluded for being pilot runs or not completing enough trials.

@@

```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Length of description produced by the teller each trial, shown on a log scale. Grey dots are individual data points, colored dots are per trial means with bootstrapped 95% CIs. \\label{length}" }
joint_data |> 
  ggplot(aes(x = trial+1, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
    geom_vline(xintercept = 4.5, lty="dashed")+
    geom_vline(xintercept = 8.5, lty="dashed")+
    geom_vline(xintercept = 12.5, lty="dashed")+
  scale_color_manual(values = time_palette) +
  geom_smooth(method = "lm", color = "black") +
  scale_y_log10()+
  xlab("Trial") +
  ylab("Words") +
  labs(color="Block")+
  facet_grid(str_c("Expt ",expt) ~ .) +
  theme(
    legend.position = "right",
    panel.grid = element_blank(),
    strip.background = element_blank()
  )
```

@@

<!-- this description metric is just the length of the description, taking out repetition/rephrasing from the child. We could also include the robustness check which is the length of all the (on-topic) teller-guesser interaction (including repeating/rephrasing, questions, I don't knows, etc), which is also very flat. -->

Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example}.

@@

```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="101%", fig.width=7, fig.height=3, fig.cap = "Semantic similarity between pairs of descriptions from different sources. Dots are means and lines are bootstrapped 95% CIs.  \\label{sim}" }

@@

@@

```{=latex}

\begin{table}
	\caption{Example descriptions children successfully used to identify different target images in Experiments 2 and 3.
}
	\label{example}
	\begin{tabular}[t]{p{14em}p{3em}}
		\hline
		%hold

		$\bullet$ vampire & \multirow{4}{*}{\includegraphics[width=.5in]{E1.jpg}}\\
		$\bullet$ a person flying &\\
		$\bullet$ a person &\\
		$\bullet$ a kite &\\
		\multicolumn{2}{l}{$\bullet$ a triangle with a head on it with feet}\\
				\multicolumn{2}{p{17em}}{$\bullet$  somebody skydiving, not in the airplane}\\
		\hline
		$\bullet$ airplane & \multirow{4}{*}{\includegraphics[width=.5in]{D1.jpg}}\\
		$\bullet$ alligator &\\
		$\bullet$ a person fell down &\\
		$\bullet$ a boat &\\
		\multicolumn{2}{p{17em}}{$\bullet$ a person that's in a race car that has one triangle and two triangles}\\
		\hline
			$\bullet$ person & 		\multirow{4}{*}{\includegraphics[width=.5in]{I1.jpg}}\\
		$\bullet$ a person holding a sandwich &\\
		$\bullet$ a people carrying a box of dirt &\\
		$\bullet$ a monster &\\
		\multicolumn{2}{p{17em}}{$\bullet$ someone holding a plate and giving it to a restaurant and has watermelon}\\
		\hline
		$\bullet$ person & \multirow{3}{*}{\includegraphics[width=.5in]{B1.jpg}}\\
		$\bullet$ a person walking &\\
		$\bullet$ a person looking down &\\
		\multicolumn{2}{p{17em}}{$\bullet$ a people, but it doesn't have any arms} \\
		\hline
	\end{tabular}
	\quad
	\begin{tabular}[t]{p{14em}p{3em}}
		\hline
		$\bullet$ a people & 		\multirow{4}{*}{\includegraphics[width=.5in]{tangram_A.png}}\\
	$\bullet$ a horse & \\
	$\bullet$ stomping & 	\\
		\multicolumn{2}{p{17em}}{$\bullet$ a person that’s walking and has a square, like on his shoulder}\\
	\multicolumn{2}{p{17em}}{$\bullet$ a diamond on the back, this leg is one leg on the ground, one leg up in the air}\\
		\hline
		$\bullet$ a bunny & \multirow{4}{*}{\includegraphics[width=.5in]{tangram_E.png}}\\
$\bullet$ a kangaroo &\\
$\bullet$ a mountain with some triangles &\\
$\bullet$ two triangles on the top &\\
\multicolumn{2}{p{17em}}{$\bullet$ a person looking backwards with a dress and a bow}\\
		\hline
		$\bullet$ a shoe &  \multirow{3}{*}{\includegraphics[width=.5in]{tangram_G.png}}\\
		$\bullet$  a walrus &\\
$\bullet$ something looking down &\\
\multicolumn{2}{p{17em}}{$\bullet$ a pterodactyl with a flag on its back }\\

\multicolumn{2}{p{17em}}{$\bullet$  a diamond on the top, and a triangle on the bottom}\\
		\hline
		$\bullet$ a bottle & \multirow{4}{*}{\includegraphics[width=.5in]{tangram_H.png}}\\
		$\bullet$ a diamond on the top &\\
 $\bullet$ a weird looking banana &\\
$\bullet$ candy &\\

		\hline
	\end{tabular}
\end{table}

@@

### Participants
Experiment 2 was run between March and August of 2024, at the same preschool as Experiment 1. No children participated in both experiments. 30 pairs of children completed all 16 critical trials, and 1 pair of children completed 10 critical trials. Our target age range was 4 and 5-year-olds, but one older 3-year-old was unintentionally included. Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months. An additional 6 games were excluded for being pilot runs or not completing enough trials.

@@

As Experiment 2 had 4 blocks, we examined whether descriptions were converging semantically toward the final description. We compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity.^[`r sim_to_last_form |> form()`] Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`). Descriptions were marginally more similar if they came from the same child, which is expected as a sign of internal consistency (`r sim_to_last_2_results |> stats_text(3,3)`). Although over time descriptions did get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant (`r sim_to_next_2_results |> stats_text(2,3)`).

@@

```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=7, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiments 2 and 3.  Heavy dots are means with bootstrapped 95% CIs. \\label{tolast}" }
joint_similarities |> 
  filter(game1==game2) |> 
  filter(later==3) |> 
  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>left_join(labels) |> 
  ggplot(aes(earlier+1, sim, color = img_source)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
    stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2), geom="line") +
    geom_smooth(color="black", method="lm")+
  facet_grid(str_c("Expt ", expt)~same_speaker) +
  scale_x_continuous(breaks=c(1,2,3), labels=c("1", "2", "3"))+
  labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
scale_color_brewer(palette="Dark2") +
  theme(
    legend.position = "right",
    legend.text = element_markdown(color = "black", size = 11),
    panel.grid = element_blank(),
    strip.background = element_blank()
  )

@@

## Methods
As Experiment 2 was very similar to Experiment 2, we focus on the changes made compared to Experiment 2. Experiment 3 was pre-registered at [anonymized link] (https://osf.io/tk7sc/overview?view_only=874df1897c2e45e187aa71944bbc2392).

@@

### Participants
Experiment 3 was run between  and YY, at the same preschool as Experiments 1 and 2. Because the stimuli changed, we allowed children who had participated in a prior experiment to participate again, and TODO participants in experiment 2 were also in experiment 3. TODO pairs of children completed all 16 critical trials, and TODO others completed ... . Our target age range was 4 and 5-year-olds. Of the TODO children, TODO were girls, and the children had a median age of TODO months, and a range of TODO-TODO months. An additional TODO games were excluded for being pilot runs, including repeat participants, not completing enough trials, or including a 3-year-old participant.

### Materials
A different set of critical images were used (see Figure \ref{game}). These were chosen from a set of commonly used tangram images [@clark1986;@hawkins2020dynamics;@boyce2024], to be fairly distinct from each other. We also switched the stimuli for the 3rd and 4th practice trials to be more abstract, so children would not be surprised by the critical stimuli not having established, canonical names.

### Procedure
The procedure was the same as for Experiment 2. <!--The bubble popping practice was modified so that both rounds of bubbles were colored dots, because children in Experiment 2 sometimes did not identify the smiley faces as more bubbles to pop. -->

@@

speed_mod_meta <- read_rds(here(mod_results, "speed_meta.rds"))
description_mod_meta <- read_rds(here(mod_results, "description_meta.rds"))

@@

As the experiments were all similar to one another, we re-ran models pooling the data across the three experiments, with experiment number as a random effect.

### Accuracy and speed 
Pooling across the three experiments, children's accuracy was above chance (OR: `r stats_text(acc_mod_meta,1)`) and accuracy numerically increased over the course of the game, although the credible interval included 0 (OR of one trial later: `r stats_text(acc_mod_meta, 2)`). Across all the experiments, children sped up over repetitions (effect of one trial later: `r stats_text(speed_mod_meta, 2)` seconds).

### Description length
Descriptions produced by tellers averaged `r stats_text(description_mod_meta, 1)` words on the first trial, and description length was relatively stable over time (change of `r stats_text(description_mod_meta, 2)` words per trial).
Utterances were more similar if they came from the same partnership (increase in cosine similarity: `r sim_across_between_mod_meta |> stats_text(2,3)`) and were more similar still if they came from the same person within the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).

### Convergence
For the other semantic analyses, we pooled across experiments 2 and 3 which had 4 blocks. In the combined model, we found that the similarity the last block utterance increased across blocks (`r stats_text(sim_to_last_mod_meta, 2,3)`) and was higher when the same child produced both utterances (`r stats_text(sim_to_last_mod_meta, 3,3)`). In a combined model, similarity to the next block utterance also increased across blocks (`r stats_text(sim_to_next_mod_meta, 2,3)`) and was higher when the same child produced both utterances (`r stats_text(sim_to_next_mod_meta, 3,3)`). Across the two experiments, utterances from different pairs to the same stimulus tended to decrease in similarity across blocks (`r stats_text(sim_across_mod_meta, 2,3)`).

@@

If children are sensitive to their partner's perspective, then whether a description was successful should influence whether the same description, or a variant, will be used in future rounds.  To test whether accuracy is predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar to the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller. This pattern of results is consistent with the expectation that children are more likely to stick to their own description or repeat the other child's description if it was previously successful. <!--Numerically, the size of the boost for being correct was similar to the size of the boost from coming from the same person `r stats_text(sim_acc_to_next_mod_meta, 6,3)`. -->

@@

Children in the late preschooler years are undergoing a time of rapidly developing linguistic and cognitive skills, so children may show age-based differences in how well they complete tasks that use language or executive functioning [TODO citation needed]. Working memory and other executive capacities are developing across childhood [@diamond2013]. Additionally, there may be gender differences at this age. Young girls have on average an advantage in early linguistic skills, showing faster earlier vocabulary growth that boys [@marchman; @rinaldi2023]. TODO needs work

@@

@@

### Accuracy 
The effects of children's age and gender on their performance are shown in Figure \ref{ageacc}. We ran a mixed effects logistic regression, where we considered possible effects of the age of each child (centered at mean age of all the children who participated) and the gender of each child on the accuracy of their selections^[`r form(acc_demog_form)`]. Accuracy was higher for older children (OR for one month older teller: `r stats_text(acc_demog_meta, 3,2)`, OR for one month older guesser: `r stats_text(acc_demog_meta, 2,2)`). Accuracy was also higher when the guesser was a girl (OR: `r stats_text(acc_demog_meta, 4,2)`), but was not credibly higher when the teller was a girl (OR: `r stats_text(acc_demog_meta, 5,2)`). The observed age and gender effects could be reflective of any of a number of developing capacities, including linguistic sophistication, improving executive functioning and impulse control, and better perspective taking and theory of mind.

### Description length
We also tested whether children's age or gender affected how many words they produced in their descriptions. We ran a mixed effects linear regression, where we considered the ages and genders of both children as possible predictors of description length^[`r form(description_demog_form)`]. Older children tended to produce longer descriptions (effect of one month older `r stats_text(description_demog_meta,3,2)`), but there was no effect of age of guesser (`r stats_text(description_demog_meta, 2,2)`). For gender, there was no effect of teller gender on description length (girls produced `r stats_text(description_demog_meta, 5,2)` words compared to boys), but tellers produced longer descriptions for boy guessers (`r stats_text(description_demog_meta, 4,2)`). There were not substantial effects of trial order or interactions between demographics and trial order.

It is curious that guesser gender but not teller gender affected description length, but for this task, it is unclear how description length would be expected to map onto linguistic sophistication for this task and demographic, since short labels could be non-specific like "person", or adequately distinguishing like "mountain".

Prominent early studies claimed that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966]. However, more recent developmental work has found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low. Recent work has established that children as young as 4 years old are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents [@leung2024]. Here, we revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.

Across 3 experiments and `r pairs` pairs of 4 and 5-year-old children, we tested how well children could produce referential expressions that allowed their partner to find a matching abstract shape. Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (`r pct`%), indicating that children were generally able to produce adequate descriptions. Additionally, children's utterances showed signs of converging toward conceptual pacts. While this task was substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.

Unlike adults, children did not display an increase in accuracy or a shortening of referential expressions over the course of the game. Still, our findings show that descriptions became increasingly similar to descriptions in the last block and that successful utterances were more similar to future utterances, suggesting that children are adapting their descriptions as the game unfolds. TODO do we want to speculate on the accuracy levels?

It is unclear to what extent the uniformly short descriptions we observed are a product of the simplified task or children's behavioral differences from adults. The low number of options and relatively distinct shapes may have obviated the need for long initial descriptions. Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020dynamics; @boyce2024]. However, young children may also struggle to produce longer descriptions, and young children may be more willing to take guesses when adults would seek additional clarification. Especially in light of other work suggesting that conceptual pact formation and reduction in utterance length sometimes decouple in adults [@boyce2024], further empirical work on the factors driving verbosity in reference games is warranted.

## Limitations
The generalizability of our results is limited by the target population, the target images, and the task structure. We sampled a convenience population of children at a university nursery school. We used two sets of 4 tangram images, which may be easier to distinguish and have higher codability than other target images used in adult iterated reference games. We specifically targeted children's abilities to construct referring expressions that can be jointly understood, so children were provided scaffolding around taking turns and talking to their partner. Thus, children's performance should be taken as a proof-of-concept about ability, rather than a claim about how generally children spontaneously demonstrate these abilities.

## Qualitative observations
Even within this population, we observed substantial pair to pair differences in what types of descriptions were used, how readily they produced them, and how well children understood and cooperated with the "rules" of the game.

@@

Occasionally, a guesser would provide descriptions of the images ("The walking person or the standing person that's holding something?") so the teller just had to indicate which one rather than self-generate a description.

@@

Other pairs of children relied on the experimenter to say whose turn it was or prompt them with options to make a guess or ask questions, and some children struggled to formulate useful questions (asking their partner just "what do you see?"). Even for children who struggled, their behaviors such as gesturing, trying to turn their iPads to show their partner, or asking the researcher for help, are indicative of attempts to solve the communication problem (albeit in ways that did not follow the "rules" of the game).

In instances where children did not communicate successfully, it is difficult to know what the limiting factor was. In general, our qualitative impression was that children were more likely to hit limits related to understanding the game structure (such as when tellers tapped their own iPad), cooperating with the game (such as guessers immediately pressing an option before the teller had a chance to say anything) or thinking the task was about correctly "naming" the shape (saying "I don't know what that's called"), than for children to specifically struggle with how to describe an unknown referent adequately.  However, pairs of children also varied significantly in how sensitive they were to collaborating with their partner and adapting descriptions based on their partners descriptions and their prior success.

By reducing the task demands and scaffolding children's interaction with each other, we were able to see that many 4-5 year old children in our sample were able to coordinate with each other on informative utterances in a repeated matching game. In a post-hoc analysis, we also observed that within this age range, older children are somewhat more successful than younger children, although there is wide individual variability. We do not claim that this ability emerges at 4 years old, although we suspect that younger children may struggle more even with the reduced task demands, making it difficult to test child-child coordination in young children.
